---
title: "The education-obesity link across countries over time"
author: "Studentnames and studentnumbers here"
date: "`r Sys.Date()`"
output: pdf_document
---

# Set-up your environment

```{r package_install, include=FALSE}
install.packages("tidyverse")
install.packages("yaml")
install.packages("rmarkdown")
```

```{r packages}
require(tidyverse)
require(rmarkdown)
require(yaml)
```

# Title Page

Include your  names
HEAD
Anastasios Goudras
Stefan
Omar
Ingmar
Diego
>>>>>>> 617f0a21df4e7799cee3e544a82591156c3d893f
Include the tutorial group number

Include your tutorial lecturer's name Simon Loop

# Part 1 - Identify a Social Problem

Use APA referencing throughout your document. [Here's a link to some explanation.](https://www.mendeley.com/guides/apa-citation-guide/)

## 1.1 Describe the Social Problem

Include the following:

-   Why is this relevant?

-   ...

# Part 2 - Data Sourcing

## 2.1 Load in the data

Both datasets are downloaded directly from Eurostat using the eurostat package, so the data is fetched when the document is knitted and nothing needs to be stored in the GitHub repository.

```{r loading_data}
library(eurostat)

# Dataset 1: obesity by education
bmi_edu <- get_eurostat("hlth_ehis_bm1e", time_format = "num")

# Dataset 2: obesity trend over time
obesity_trend <- get_eurostat("sdg_02_10", time_format = "num")
```

## 2.2 Provide a short summary of the dataset(s)

```{r}
head(bmi_edu)
head(obesity_trend)
```
Our first dataset, bmi_edu, comes from Eurostat and is called "Body mass index (BMI) by sex, age and educational attainment level" (dataset code hlth_ehis_bm1e). It is based on the European Health Interview Survey (EHIS), a survey in which people across European countries report their own height and weight. For each country and year, the dataset gives the percentage of people in different BMI categories (such as normal weight, overweight, and obese), broken down by education level, sex, and age. Because the survey is only carried out every few years, the data covers a small number of waves (around 2014 and 2019). This dataset is central to our project because it is the one that connects weight to education level.

Our second dataset, obesity_trend, also comes from Eurostat and is called "Obesity rate by body mass index (BMI)" (dataset code sdg_02_10). It is one of Eurostat's Sustainable Development Goal (SDG) monitoring indicators, built from European health surveys in which adults report their own height and weight. For each country and year, it gives the percentage of people in different BMI categories such as overweight and obese. Unlike our first dataset, it does not break the figures down by education level, but it covers more years per country (for example 2008, 2014, 2017, 2019 and 2022), which makes it well suited for showing how obesity has changed over time. This dataset is important for our project because it provides the "over time" and "across countries" side of our research question.

## 2.3 Describe the type of variables included

The variables in our datasets fall into a few types. The main health information is the BMI/obesity measure — the share of people who are overweight or obese (the bmi and values columns). The main socioeconomic (SES) information is the education level (the isced11 column in bmi_edu), which is the variable at the centre of our research question. The remaining variables — sex, age, country (geo) and year (TIME_PERIOD) — are background characteristics that let us split the figures by group, country and time.
In terms of how the data was collected, neither dataset comes from administrative sources such as hospital or government records. Both are based on surveys in which people are interviewed and report their own height and weight. This has two consequences. First, because height and weight are self-reported, obesity tends to be slightly underestimated, as people often round their weight down. Second, our two datasets are not fully independent, because the SDG indicator (obesity_trend) is partly built on the same European Health Interview Survey as bmi_edu — so they are related rather than two completely separate sources.
We chose these two datasets together on purpose: bmi_edu provides the education breakdown we need, while obesity_trend gives a longer view over time and across countries. Because both rely on self-reported survey data, our analysis can describe patterns and associations — for example, that obesity is more common among lower-educated groups — but it cannot prove that education causes obesity.

# Part 3 - Quantifying

## 3.1 Data cleaning

Say we want to include only larger distances (above 2) in our dataset, we can filter for this.

```{r data_cleaning}
mean(dataset$percollege)
```

Please use a separate 'R block' of code for each type of cleaning. So, e.g. one for missing values, a new one for removing unnecessary variables etc.

## 3.2 Generate necessary variables

Variable 1

```{r gen_var1}

```

Variable 2

```{r gen_var2}

```

## 3.3 Visualize temporal variation

```{r}

```

## 3.4 Visualize spatial variation

```{r visualise_map}

```

Here you provide a description of why the plot above is relevant to your specific social problem.

## 3.5 Visualize sub-population variation

What is the poverty rate by state?

```{r visualise_subpopulations}
dataset$inmetro <- dataset$inmetro %>% as.factor()
# Boxplot of poverty rate by state using the 'midwest' dataset
ggplot(dataset, aes(x = inmetro, y = percadultpoverty)) +
  geom_boxplot() +
  labs(
    title = "Distribution of Poverty Rates by Metropolitan status (Midwest counties)",
    x = "Metropolitan Area",
    y = "Poverty Rate of Adults (%)"
  ) +
  theme_minimal() +
  theme(
    legend.position = "right"
  )
```

Here you provide a description of why the plot above is relevant to your specific social problem.

## 3.6 Event analysis

Analyze the relationship between two variables.

```{r analysis}

```

Here you provide a description of why the plot above is relevant to your specific social problem.

# Part 4 - Discussion

## 4.1 Discuss your findings

# Part 5 - Reproducibility

## 5.1 Github repository link

Provide the link to your PUBLIC repository here: ...

## 5.2 Reference list

Use APA referencing throughout your document.
